---
title: "141124test"
author: "Yakovlev D.M."
date: "2024-11-14"
output: html_document
---

### 6-1. Civil Marriage Act in Canada (2005)

Analyze the results with a significance test, stating any assumptions, and interpret the p-value
```{r}
# According to 2017 survey. p - percentage of supporters, n - number of Canadians
n<-3402
p<-0.73
```
Assuming we want to know whether the majority or minority supports the act, we should set null hypothesis:

H0: p<=0.5 (assuming that the minority supports the act) 

H1: p>0.5 (the majority supports the act)

We will set alpha as 0.10
```{r}
alpha<-.10
```
Let us use one sided ("less" for the alternative) proportion test and interpret the results.
```{r}
res<-prop.test(n*p, n, alternative="greater")
res
res$p.value
```
As we may notice, p-value ~ 0 < alpha = 0.1, which means that the alternative hypothesis (H1) will be accepted. So, we may assume that the majority supports the act.


### 6-2. 2018 General Social Survey.

Test that the population proportions are equal for females and males. Report and interpret the p-value
```{r}
# Survey results
female_total<-1178
female_yes<-1017
p_female<-female_yes/female_total
male_total<-945
male_yes<-703
p_male<-male_yes/male_total
```

In order to test the population proportions we shall use prop.test. Let us arrange null and alternative hypothesis

H0: p_male = p_female (proportions of male and female opinions are equal)
H1: p_male != p_female (proportions are unequal)

We shall use alpha = 0.10
```{r}
alpha <- 0.10
```

```{r}
res<-prop.test(c(female_yes, male_yes), c(female_total, male_total))
res
res$p.value
```
p-value ~ 4.6*10^-12 < alpha = 0.10, which means that we reject the H0, and therefore, accept the H1. As p-value is unsignificant, we may suggest that female and male beliefs on life and death are definitely different from one another on average.

### 6-3. Automobile dealership

To collect data on the plan, a sample of sales personnel will be allowed to
sell under the new bonus plan for a one-month period.

Three points are to be made:
a. Develop the null and alternative hypotheses most appropriate for this situation.
b. Comment on the conclusion when H0 cannot be rejected
Ñ. Comment on the conclusion when H0 can be rejected
```{r}
# Mean sales volume (msv, automobiles/month)
msv<-14
```

a. Since we are planning on increasing the sales volume, we should conduct hypothesis tests for the mean. We shall use t-test in this instance. We will set null and alternative hypotheses

Assume we have a sample with new mean sales volume
```{r}
msv_new<-21
```
H0: msv_new <= msv (New bonus plan garnered less sales) 

H1: msv_new > msv (New bonus plan garnered more sales)

b.
H0 cannot be rejected if we cannot say that the new bonus plan yields more sales, that is when msv_new is close to or less than msv.
```{r}

```

c.
H0 can be rejected if we can definitely say that the new bonus plan yields more sales, that is when msv_new is significantly larger that msv. 
```{r}

```


### 6-4. Federal income tax

```{r}
# average refund ($)
average_refund<-1056
```

Four points are to be discussed:
a. Develop appropriate hypotheses.
Some people believe that getting a large tax refund is not as desirable as more accurate withholding throughout the year, as a large refund represents a loan paid back by the government interest-free. Therefore, individuals attempt to lower their tax refunds. 

Let us construct null and alternative hypotheses

H0: avr_refund_last_min >= average_refund (the average refund when filing at the last five days is greater than in general) 

H1: avr_refund_last_min < average_refund (assuming that average refund during the last five days is lower than the total average)

```{r}

```
b. What is p-value?
```{r}
# sample of last-minute filers
n<-400
# sample mean
avr_refund_last_min<-910
# standard deviation
sd<-1600
```

```{r}
z<-sqrt(n)*(avr_refund_last_min-average_refund)/sd
z
```
P-value
```{r}
p_val<-pnorm(z)
p_val
```
c. At alpha = .05, what is your conclusion?
```{r}
alpha<-.05
alpha<p_val
```
Since p-value < alpha, we may conclude that H0 can be rejected, and therefore, H1 is accepted, meaning that the last minute filers are likely to have less refund on average compared to a total average.

d. Critical value approach
```{r}
z
```

```{r}
crit_val<-qnorm(alpha)
crit_val
```

```{r}
z<crit_val
```

Since z-value is less than critical value, z-value is inside of critical region, so we shall reject null hypothesis.

### 6-5. How Undergraduate Students Use Credit Cards (Sallie Mae, April 2009)

```{r}
# mean credit card balance (mccb, $)
mccb<-3173
# standard deviation ($)
sd<-1000
p<-0.44
```

a. State the null and alternative hypotheses

Since we are assured of all-time high mean credit card balance, we are more prone to assume that this tendency shall continue. Therefore, let us construct null and alternative hypotheses.

H0: mccb_new<=mccb (the contemporary average is lesser than it was)

H1: mccb_new>mccb (the contemporary average is greater than it was)
```{r}

```
b. P-value if sample = 180, sample mean = 3325$?
```{r}
n<-180
sample_mean<-3325 #mccb_new
```

```{r}
z<-sqrt(n)*(sample_mean-mccb)/sd
z
```

```{r}
p_val<-1-pnorm(z)
p_val
```
c. alpha = 0.05, what is the conclusion?
```{r}
alpha<-0.05
p_val<alpha
```

p-value ~ 0.021 < alpha = 0.05 which concludes that we reject H0 and accept H1. It could be interpreted that there is evidence of ever-increasing tendency regarding mean credit card balance for undergraduate students.

### 6-6. U.S. Diversified Equity (BusinessWeek, January 26, 2004).

```{r}
# Average annual total return
p <- 0.041
```

3 points:
a. Formulate the hypotheses to determine whether the mean annual return for mid-cap growth funds differ from the mean for U.S. Diversified Equity funds.

As we put our interest into finding the difference between mid-cap growth funds and U.S. Diversified Equity funds, we should construct the following null and alternative hypotheses:

H0: mc_p = p (mid-cap growth funds' average is alike U.S. Diversified Equities funds)

H1: mc_p != p (the funds are significantly different)
```{r}

```
b. sample - 40, x_mean = 0.034, sd = 0.2
```{r}
n<-40
x_mean<-0.034 #mc_p
sd<-0.2
```

```{r}
z<-sqrt(n)*(x_mean - p)/sd
z
```

```{r}
p_val<-2*(1-pnorm(abs(z)))
p_val
```
c. alpha = 0.05, conclusions.
```{r}
alpha<-0.05
p_val>alpha
```

Since p_val ~ 0.825 > alpha = 0.05, we cannot reject H0, which means that we cannot deem the difference between mid-cap growth funds and U.S. Diversified Equity funds as significantly different.


### 6-7. Shareholders' group (The Wall Street Journal, January 2, 2007)

The claim is that the mean tenure for a CEO was at least nine years.
```{r}
# claimed mean tenure
claim <- 9
# sample mean tenure for a chief executive office (CEO, in years)
x_mean <- 7.27

# standard deviation (in years)
s<-6.38
```
a. Formulate hypotheses to challenge the validity of the claim made by the shareholders' group.

To construct null and alternative hypotheses, we have to mind the claim of the supposed mean tenure for a CEO, which will be used.

H0: claim >= 9 (the claim is true and is at least nine years)

H1: claim < 9 (assuming that the claim is false)
```{r}

```
b. Assume 85 were included in the sample. What's the p-value?
```{r}
n<-85
z<-sqrt(n)*(x_mean - claim)/s
z
```
```{r}
p_val<-pnorm(z)
p_val
```
c. alpha = 0.01, conclusion.
```{r}
alpha<-0.01
p_val<alpha
```

Since p_val ~ 0.006 < alpha = 0.01, we can reject H0 and accept H1, which mean there is evidence to tell that the mean tenure for a CEO was less than nine years.


### 6-8. Shorney Construction Company

```{r}
# sample size
n<-30
# standard deviation (minutes/day)
sd<-20
# mit
mit <- 72
```
It is assumed that mean idle time (mit) per worker is 72 or fewer minutes per day.

a. Hypotheses
H0: mit <= 72 (assuming that workers have less than 72 minutes of idle time per day)

H1: mit > 72 (the workers have more idle time than assumed)

b. Probability of Type II error if mit = 80?
Considering alpha = 0.05
```{r}
alpha<-0.05
```

```{r}
z_alpha<-qnorm(alpha, lower.tail = FALSE)
z_alpha
```

```{r}
mit_new<-80
shift<-sqrt(n)*(mit_new-mit)/sd
typeIIerror<-pnorm(z_alpha, mean=shift)
typeIIerror
```
Type II error ~ 0.2925
c. -||- if mit = 75?
```{r}
mit_new<-75
shift<-sqrt(n)*(mit_new-mit)/sd
typeIIerror<-pnorm(z_alpha, mean=shift)
typeIIerror
```
Type II error ~ 0.7948
d. -||- if mit = 70?
```{r}
mit_new<-70
shift<-sqrt(n)*(mit_new-mit)/sd
typeIIerror<-pnorm(z_alpha, mean=shift)
typeIIerror
```

Type II error ~ 0.9858

### 6-9. Joan's Nursery

The claim is that the mean tree-planting time differs from two hours
```{r}
# sample (hours/plant)
sample<-c(1.7, 1.5, 2.6, 2.2, 2.4, 2.3, 2.6, 3.0, 1.4, 2.3)
n<-length(sample)
n
alpha<-0.05
```

4 points:
a. State the null and alternative hypotheses

Since managers use two-hours system for the planting of a medium-sized tree to appraise the labor cost, we may want to construct hypotheses that make sure the two-hours system is intact.

However, if we, instead, show that the actual mean time is lower than two hours, it would increase the estimated labor cost which would be profitable.

H0: mean >= 2 (mean time is greater than 2 hours)

H1: mean < 2 (mean time is less than 2 hours)

b. Compute the sample mean
```{r}
smp_mean<-mean(sample)
smp_mean
```
c. Compute the sample standard deviation
```{r}
smp_sd<-sd(sample)
smp_sd
```
d. p-value and conclusion
```{r}
z<-sqrt(n)*(smp_mean-2)/sd
z
```

```{r}
p_val<-pnorm(z)
p_val
```

p-value ~ 0.5126 > alpha = 0.05. Therefore, we cannot reject H0, so it is unlikely that the mean labor time is less than 2 hours.


### 6-10. Consumer Reports

```{r}
# percentage of shoppers that say that supermarket brands are as good as national name brands.
p<-0.64
```

a. Formulate hypotheses
Ketchup problem. It is assumed that, since the national name-brand manufacturer has made an investigation, it is in the researchers interest to show the results that would please the manufacturer. Let us construct the hypotheses

H0: p >= 0.64 (shoppers deem supermarket and national brand ketchup as equals for the most part)

H1: p < 0.64 (shoppers are less likely to deem supermarket and national brand ketchup as equals)

```{r}

```
b. sample = 100, both is good = 52, p-value
```{r}
# sample size
n<-100
# both is good
big<-52
p_mean<-big/n
```

```{r}
sd<-sqrt(p*(1-p))
z<-sqrt(n)*(p_mean-p)/sd
z
```

```{r}
res<-binom.test(big, n, p, alternative="less")
res
```
c. alpha = 0.05, conclusion.
```{r}
alpha<-0.05
res$p.value>alpha
```
Since the p-value ~ 0.009 < alpha = 0.05, it means that we can reject H0 and accept H1. Therefore, shoppers are not as inclined to deem supermarket ketchup and national name brand ketchup as equals as it was shown by Consumer Reports.

d. National brand ketchup manufacturers pleased with the conclusion?

In my humble opinion, they would be pleased to know that less shoppers prefer both supermarket and national brand ketchups than was previously stated, which yields better ketchup sales for national name-brand manufacturer.
```{r}

```


### 6-11. Consumer Research Group.

Claim [from automobile manufacturer] that a new economy model will travel at least 25 miles per gallon of gasoline.
```{r}
mean_mpg<-25
```
a. alpha = 0.02, sample = 30 (cars), sd = 3 miles/gallon. Rejection rule based on x_mean.

So, we are interesting in testing an automobile manufacturer's claim.
We will make rejection rule based on the value of x_mean in this manner:

H0: mean_mpg >= 25 (the manufacturer's claim)

H1: mean_mpg < 25 (worse-than-claimed performance)

```{r}
alpha<-0.02
sample<-30
sd<-3
```

For instance, assume that x_mean = 28 mpg
```{r}
x_mean <- 24.3
```

```{r}
z<-sqrt(sample)*(x_mean-mean_mpg)/sd
z
```

```{r}
p_val<-pnorm(z)
p_val
```

```{r}
p_val > alpha
```
So, we propose the rejection rule based on z-score and p-value. If p-value is greater than alpha, we cannot reject H0, otherwise we can reject H0 and accept H1.

b. Type II error if the actual mileage is 23 miles/gallon
```{r}
z_alpha<-qnorm(alpha)
z_alpha
```

```{r}
mpg_new<-23
shift<-sqrt(sample)*(mpg_new-mean_mpg)/sd
shift
typeIIerror<-pnorm(z_alpha, mean=shift, lower.tail=FALSE)
typeIIerror
```
Type II error ~ 0.055
c. -||- 24 mpg
```{r}
mpg_new<-24
shift<-sqrt(sample)*(mpg_new-mean_mpg)/sd
shift
typeIIerror<-pnorm(z_alpha, mean=shift, lower.tail=FALSE)
typeIIerror
```
Type II error ~ 0.5902
d. -||- 25.5 mpg
```{r}
mpg_new<-25.5
shift<-sqrt(sample)*(mpg_new-mean_mpg)/sd
shift
typeIIerror<-pnorm(z_alpha, mean=shift, lower.tail=FALSE)
typeIIerror
```
Type II error ~ 0.9985