---
title: "collegeFirst"
author: "Yakovlev D.M."
date: "2024-11-08"
output: html_document
---

# Ђнализ датасета

```{r}
library(tidyverse)
library(GGally)
library(ggplot2)
```

### 0. РџСЂРµРґРІР°СЂРёС‚РµР»СЊРЅС‹Р№ Р°РЅР°Р»РёР· РґР°РЅРЅС‹С…
```{r}
library(readxl)
col_I_sn<-read_excel("stat2021_sm\\us_col.std\\I_shortname.xls")
col_I<-read_excel("stat2021_sm\\us_col.std\\I.xls")
col_I
```

```{r}
summary(college_I)
```

```{r}
names(college_I)
```

```{r}
col_I_sn$NEW10
```
РљРѕР»РёС‡РµСЃС‚РІРµРЅРЅС‹Рµ РїСЂРёР·РЅР°РєРё:

* AVRMATH Average Math SAT score                     
* AVRVERB Average Verbal SAT score                   
* AVRCOMB Average Combined SAT score
* AVR_ACT Average ACT score                     
* MATH_1 First quartile - Math SAT                    
* MATH_3 Third quartile - Math SAT                   
* VERB_1 First quartile - Verbal SAT
* VERB_3 Third quartile - Verbal SAT
* ACT_1 First quartile - ACT                          
* ACT_3 Third quartile - ACT              
* APP_REC Number of applications received             
* APP_ACC Number of applicants accepted             
* NEW_STUD Number of new students enrolled   
* FULLTIME Number of fulltime undergraduates          
* PARTTIME Number of parttime undergraduates
* IN_STATE In-state tuition                        
* OUT_STAT Out-of-state tuition                       
* R_B_COST Room and board costs
* ROOM Room costs                                    
* BOARD Board costs                              
* ADD_FEE Additional fees                            
* BOOK Estimated book costs                 
* PERSONAL Estimated personal spending
* PH_D Pct. of faculty with Ph.D.'s          
* TERM_D Pct. of faculty with terminal degree         
* SAL_FULL Average salary - full professor          
* SAL_AC Average salary - associate professor         
* SAL_AS Average salary - assistant professor         
* SAL_ALL Average salary - all ranks       
* COMP_FUL Average compensation - full professor   
* COMP_AC Average compensation - associate professor  
* COMP_AS Average compensation - assistant professor  
* COMP_ALL Average compensation - all ranks           
* NUM_FULL Number of full professor
* NUM_AC Number of associate professor                
* NUM_AS Number of assistant professor                
* NUM_INS Number of instructors                
* NUM_ALL Number of faculty - all ranks   
* INSTRUCT Instructional expenditure per student
* GRADUAT Graduation rate
* SF_RATIO Student/faculty ratio
* DONATE Pct.alumni who donate
* NEW10 Pct. new students from top 10% of H.S. class - % студентов из топ 10% своей старшей школы
* NEW25 Pct. new students from top 25% of H.S. class - % студентов мз топ 25% своей старшей школы
(‹ибо написать - всЮ остальное)

ЏорЯдковые признаки:

* FICE - Federal ID Number

Љачественные признаки:

* ...1 - Ќазвание университета 
* PPIND Public/private indicator (public=1, private=2)
* STATE State (postal code)
* TYPE - I

```{r}
head(col_I_sn)
summary(col_I_sn)
```

(Љомментарий убрать:) ‚ данных есть пропуски. ‹ибо заполнЯем их (если знаем, как), либо оставлЯем как есть.

```{r}
dim(col_I_sn)
```

```{r}
col_I_sn$TYPE
```
TYPE состоит из одних и тех же значений, поэтому его можно убрать

```{r}
col_I_sn<-col_I_sn|>select(-TYPE)
```

```{r}
names(col_I_sn)
```

Љолледжи делЯтсЯ на государственные (public) и частные (private). ђазделим датасет на две части.

```{r}
col_prv<-col_I_sn|>filter(PPIND==2)
head(col_prv)
```

```{r}
col_pub<-col_I_sn|>filter(PPIND==1)
head(col_pub)
```

```{r}
names_col<-names(col_I_sn)
names_col
avr_col<-names_col[grep("AVR", names_col)]
avr_col
act_col<-names_col[grep("ACT", names_col)]
num_col<-names_col[grep("NUM", names_col)]
sal_col<-names_col[grep("SAL", names_col)]
comp_col<-names_col[grep("COMP", names_col)]
new_col<-names_col[grep("NEW", names_col)]
math_col<-names_col[grep("MATH", names_col)]
verb_col<-names_col[grep("VERB", names_col)]
app_col<-names_col[grep("APP", names_col)]
time_col<-names_col[grep("TIME", names_col)]
cat_col<-c("PPIND", "FICE", "STATE")
stat_col<-c("IN_STATE", "OUT_STAT")
misc_col<-c("R_B_COST", "ROOM", "BOARD", "ADD_FEE", "BOOK", "PERSONAL", "INSTRUCT")
```

### Private
ђассмотрим частные колледжи.
```{r}
ggpairs(col_prv[c(sal_col)], aes(alpha = 0.5),
        diag = list(continuous = wrap("barDiag", bins = 20)))
```

```{r}
ggpairs(col_prv[c(comp_col)], aes(alpha = 0.5),
        diag = list(continuous = wrap("barDiag", bins = 20)))
```

```{r}
ggpairs(col_prv[c(num_col)], aes(alpha = 0.5),
        diag = list(continuous = wrap("barDiag", bins = 20)))
```

```{r}
ggpairs(col_prv[c(act_col)], aes(alpha = 0.5),
        diag = list(continuous = wrap("barDiag", bins = 20)))
```

```{r}
ggpairs(col_prv[c(avr_col)], aes(alpha = 0.5),
        diag = list(continuous = wrap("barDiag", bins = 20)))
```

```{r}
ggpairs(col_prv[c(stat_col, misc_col)], aes(alpha = 0.5),
        diag = list(continuous = wrap("barDiag", bins = 20)))
```
```{r}
ggpairs(col_prv[c(app_col, time_col)], aes(alpha = 0.5),
        diag = list(continuous = wrap("barDiag", bins = 20)))
```

```{r}
ggpairs(log(col_prv[c(app_col, time_col)]), aes(alpha = 0.5),
        diag = list(continuous = wrap("barDiag", bins = 20)))
```

```{r}
sm<-lm(OUT_STAT~IN_STATE+ADD_FEE+0, col_I_sn)
summary(sm)
```
### Public
овалвалпжфлщцтмтчмюищуацхщплше
### 1. ‚иды распределений и сравнение распределений
```{r}
library(ggpubr)
```

### Private
```{r}

```